---
title: "More complicated graphs"
author: "Fiona Messer"
date: "2026-09-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
source("ggplot2_theme_FM.R")
```

I want to look at a region and several countries at the same time. I could go back to the original data set before I split it up, but I want to keep the columns I added to `alcohol_country` which give the parent WHO region for each country.

Instead, I will combine the `alcohol_country` and `alcohol_whoregion` data frames.

```{r Combine country and WHO region dfs}
alcohol_reg_country <- bind_rows(alcohol_country, alcohol_whoregion)
alcohol_reg_country$X <- NULL
```

Next I will plot alcohol consumption over time for Europe and five random European countries.

## Alcohol consumption in Europe between 2000-2022

First I'll pick five countries at random:

```{r Country picker}
country <- filter(WHO_REGIONS, ParentLocation == "Europe")
country <- sample(country$GEO_NAME_SHORT, size = 5, replace = FALSE)
```

Next plot just the data for the region of Europe on a ribbon plot:

```{r Europe alcohol consumption}
alcohol_reg_country %>% 
  filter(alcohol_reg_country$GEO_NAME_SHORT == "Europe") %>%
  ggplot(aes(x=DIM_TIME, y = RATE_PER_CAPITA_N, ymin = RATE_PER_CAPITA_NL, ymax = RATE_PER_CAPITA_NU)) +
  geom_ribbon(alpha = 0.25, colour = "#BC5F04", fill = "#BC5F04") +
  geom_line(colour = "#010001") +
  theme_FM() +
  labs(title = paste0("Alcohol consumption rate of adults in Europe"), 
       subtitle = "Litres per year per capita",
       x = NULL,
       y = "Rate (L/Year) per Capita") +
  xlim(2000, 2022) +
  ylim(0, max(alcohol_reg_country$RATE_PER_CAPITA_NU))
```

Next I'll plot the five random countries I picked earlier. I am going to make a new object, `plotting_data`, to make it easier to handle which parts of the data set I'm trying to plot at any given moment.

```{r plotting_data}
if (nrow(plotting_data) > 0) {
  plotting_data <- NULL
} 
plotting_data <- data.frame()
for (c in country) {
  plotting_data <- rbind(filter(alcohol_reg_country, GEO_NAME_SHORT == c), plotting_data)
}
```

I have the data organised, next I'll make the plot.

```{r Five European countries}
ggplot(plotting_data, 
       aes(x = DIM_TIME, 
           y = RATE_PER_CAPITA_N,
           color = GEO_NAME_SHORT)) +
  geom_line() +
  scale_color_manual(values = autumn) + # "autumn" is a pre-defined color palette in my source file
  theme_FM() +
  labs(title = paste0("Alcohol consumption rate of adults in Europe"), 
       subtitle = "Litres per year per capita",
       x = NULL,
       y = "Rate (L/Year) per Capita",
       color = "Country") +
  xlim(2000, 2022) +
  ylim(0, (max(plotting_data$RATE_PER_CAPITA_NU)*1.05)) #This sets the y axis maximum to 5% more than the largest upper limit in the data
```

Now I'll add the data for the whole region of Europe:

```{r add Europe to plotting data}
#plotting_data <- rbind(filter(alcohol_reg_country, GEO_NAME_SHORT == "Europe"), plotting_data)

ggplot(plotting_data, 
       aes(x = DIM_TIME)) +
  geom_line(aes(y = RATE_PER_CAPITA_N,
            color = GEO_NAME_SHORT)) +
  scale_color_manual(values = dusty) + # "dusty" is a pre-defined color palette in my source file
  theme_FM() +
  labs(title = paste0("Alcohol consumption rate of adults in Europe"), 
       subtitle = "Litres per year per capita",
       x = NULL,
       y = "Rate (L/Year) per Capita",
       color = "Country") +
  xlim(2000, 2022) +
  ylim(0, (max(plotting_data$RATE_PER_CAPITA_NU)*1.05)) #This sets the y axis maximum to 5% more than the largest upper limit in the data
```

