---
title: "data_import"
author: "Fiona Messer"
date: "2026-08-28"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data
Load the dataframe downloaded from WHO 
https://data.who.int/indicators/i/EF38E6A/EE6F72A
```{r}
alcohol <- read.csv("alcohol_consumption.csv")
summary(alcohol)
head(alcohol)
```

## Simplify the dataframe

There are a few columns which we don't need to keep, for example the first four columns contain the same code for each line of the data.

```{r}
alcohol <- alcohol[,5:15]
head(alcohol)
```

The columns `DIM_TIME_TYPE`, `DIM_PUBLISH_STATE_CODE`, `IND_NAME` and `DIM_SEX` are also all the same for all of the data as well, so they can be removed.

```{r}
alcohol$DIM_TIME_TYPE <- NULL
alcohol$DIM_PUBLISH_STATE_CODE <- NULL
alcohol$IND_NAME <- NULL
alcohol$DIM_SEX <- NULL
head(alcohol)
```

Now we only have the interesting columns of the dataframe, so we can get on with playing with the data.

## Sort the data into new datframes

Looking at the variable `DIM_GEO_CODE_TYPE`, I can see that there are several different geographical levels/classifications in this column.

```{r}

```

