---
title: "data_import"
author: "Fiona Messer"
date: "2026-08-28"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data
Load the data frame downloaded from WHO 
https://data.who.int/indicators/i/EF38E6A/EE6F72A
```{r}
alcohol <- read.csv("alcohol_consumption.csv")
summary(alcohol)
head(alcohol)
```

## Simplify the dataframe

There are a few columns which we don't need to keep, for example the first four columns contain the same code for each line of the data.

```{r}
alcohol <- alcohol[,5:15]
head(alcohol)
```

The columns `DIM_TIME_TYPE`, `DIM_PUBLISH_STATE_CODE`, `IND_NAME` and `DIM_SEX` are also all the same for all of the data as well, so they can be removed.

```{r}
alcohol$DIM_TIME_TYPE <- NULL
alcohol$DIM_PUBLISH_STATE_CODE <- NULL
alcohol$IND_NAME <- NULL
alcohol$DIM_SEX <- NULL
head(alcohol)
```

Now we only have the interesting columns of the data frame, so we can get on with playing with the data.

## Sort the data into new datframes

Looking at the variable `DIM_GEO_CODE_TYPE`, I can see that there are several different geographical levels/classifications in this column.

```{r}
unique(alcohol$DIM_GEO_CODE_TYPE)
table(alcohol$DIM_GEO_CODE_TYPE)
```

I will separate out the data into new data frames based on the `DIM_GEO_CODE_TYPE` using the `dplyr` package.

```{r}
library(dplyr)
alcohol_class <- filter(alcohol, DIM_GEO_CODE_TYPE == "CLASSIFICATION")
alcohol_country <- filter(alcohol, DIM_GEO_CODE_TYPE == "COUNTRY")
alcohol_global <- filter(alcohol, DIM_GEO_CODE_TYPE == "GLOBAL")
alcohol_region <- filter(alcohol, DIM_GEO_CODE_TYPE == "REGION")
alcohol_whoregion <- filter(alcohol, DIM_GEO_CODE_TYPE == "WHOREGION")
```

These data sets are now set up to use for making graphs.

## Alternative dataset

The WHO also has another data set, which further subdivides alcohol consumption by beverage type.
https://www.who.int/data/gho/data/indicators/indicator-details/GHO/alcohol-recorded-per-capita-(15-)-consumption-(in-litres-of-pure-alcohol)

```{r}
beverage_type <- read.csv("alcohol_bev_type.csv")
head(beverage_type)
summary(beverage_type)
```

Like the previous data set, there are a few columns which contain the same information for each entry, so they will be removed.

```{r}
beverage_type <- beverage_type[, c(4, 5, 7, 8, 10, 11, 13, 14, 21, 22, 24, 30, 31, 33, 34)]
head(beverage_type)
```

## Regions and parent locations

Both data sets refer to regions or classifications - these are groups of countries based on factors such as geographical location or income.

The `beverage_type` data set makes it obvious which country belongs to which 'parent region' with the `ParentLocation` column. The `alcohol` data set does not have this built in, so it would be useful to have a grouping reference for when I want to compare between levels.

I'll retrieve the  parent regions from `beverage_type` to see if they match up with the ones from `alcohol`:

```{r}
unique(beverage_type$ParentLocation)
unique(alcohol_whoregion$GEO_NAME_SHORT)
```

As we can see, `ParentLocation` in `beverage_type` and `whoregion` in `alcohol` use the same classification system. So we can extract the classification from the `beverage_type` data set, and apply it to the alcohol data set.

```{r}
WHO_REGIONS <- beverage_type[, 1:4]
WHO_REGIONS <- (WHO_REGIONS %>% distinct(Location, .keep_all = TRUE))
```

`WHO_REGIONS` is now a reference table available in the **data_import_files** folder.

I can add the WHO Region back into the `alcohol_country` data set.

```{r}
WHO_REGIONS$GEO_NAME_SHORT <- WHO_REGIONS$Location
WHO_REGIONS$Location <- NULL
alcohol_country <- merge(WHO_REGIONS, alcohol_country, by = "GEO_NAME_SHORT")
```

## Classifications

The alcohol data set has data on `Classification`, which refers to a UN classification system of development https://unstats.un.org/unsd/methodology/m49/

```{r}
unique(alcohol_class$GEO_NAME_SHORT)
```

I might add this into the `alcohol_country` data set later, but for now I will leave it out.
