---
title: "Simple graphs"
author: "Fiona Messer"
date: "2026-09-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## Making simple plots

Now the data sets are nicely formatted I can start messing around with plotting the data in fun and interesting ways.

Load the data from the `data_import_files` folder:

```{r}
alcohol_country <- read.csv("alcohol_country.csv")
```

I will start by plotting the alcohol consumption data for a random country over time. First, choose a country:

```{r}
country <- unique(alcohol_country$GEO_NAME_SHORT)[[sample(1:length(unique(alcohol_country$GEO_NAME_SHORT)), 1)]]
filter(alcohol_country, GEO_NAME_SHORT == country)
print(country)
```

Next, plot a scatterplot with a line:

```{r}
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME, y=RATE_PER_CAPITA_N)) +
    geom_line() +
    geom_point()
```

Plot made, can I add the upper and lower estimates as well?

```{r}
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME)) +
    geom_line(aes(y = RATE_PER_CAPITA_N, colour = "Rate per capita")) +
    geom_line(aes(y = RATE_PER_CAPITA_NL, colour = "Lower limit")) +
    geom_line(aes(y = RATE_PER_CAPITA_NU, colour = "Upper limit"))
```

Alternatively with error bars:

```{r}
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME)) +
  geom_line(aes(y = RATE_PER_CAPITA_N)) +
  geom_errorbar(aes(ymin = RATE_PER_CAPITA_NL, ymax = RATE_PER_CAPITA_NU))
  
```

Which looks ugly.
