---
title: "Simple graphs"
author: "Fiona Messer"
date: "2026-09-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## Making simple plots

Now the data sets are nicely formatted I can start messing around with plotting the data in fun and interesting ways.

Load the data from the `data_import_files` folder:

```{r read csv}
alcohol_country <- read.csv("alcohol_country.csv")
```

I will start by plotting the alcohol consumption data for a random country over time. First, choose a country:

```{r Choose country}
country <- unique(alcohol_country$GEO_NAME_SHORT)[[sample(1:length(unique(alcohol_country$GEO_NAME_SHORT)), 1)]]
filter(alcohol_country, GEO_NAME_SHORT == country)
print(country)
```

Next, plot a scatterplot with a line:

```{r Scatter and line plot}
#| fig.cap: !expr paste0("Alcohol consuption rate (L/year) per capita, ", country)
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME, y=RATE_PER_CAPITA_N)) +
    geom_line() +
    geom_point()
```

Plot made, can I add the upper and lower estimates as well?

```{r Upper and lower limits}
#| fig.cap: !expr paste0("Alcohol consuption rate (L/year) per capita, ", country, ", with errors")
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME)) +
    geom_line(aes(y = RATE_PER_CAPITA_N, colour = "Rate per capita")) +
    geom_line(aes(y = RATE_PER_CAPITA_NL, colour = "Lower limit")) +
    geom_line(aes(y = RATE_PER_CAPITA_NU, colour = "Upper limit"))
```

Alternatively with error bars:

```{r Error bars}
#| fig.cap: !expr paste0("Alcohol consuption rate (L/year) per capita, ", country, ", with error bars")
alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME)) +
  geom_line(aes(y = RATE_PER_CAPITA_N)) +
  geom_errorbar(aes(ymin = RATE_PER_CAPITA_NL, ymax = RATE_PER_CAPITA_NU))
  
```

Which looks ugly. Try the `geom_ribbon()` function instead.

```{r Error ribbon}
#| fig.cap: !expr paste0("Alcohol consuption rate (L/year) per capita, ", country, ", with error ribbon")

source("ggplot2_theme_FM.R")

alcohol_country %>% 
  filter(alcohol_country$GEO_NAME_SHORT == country) %>%
  ggplot(aes(x=DIM_TIME, y = RATE_PER_CAPITA_N, ymin = RATE_PER_CAPITA_NL, ymax = RATE_PER_CAPITA_NU)) +
  geom_ribbon(alpha = 0.25, colour = "#BC5F04", fill = "#BC5F04") +
  geom_line(colour = "#010001") +
  theme_FM() +
  labs(title = paste0("Alcohol consumption rate of adults in ", country), 
       subtitle = "Litres per year per capita",
       x = NULL,
       y = "Rate (L/Year) per Capita") +
  xlim(2000, 2022) +
  ylim(0, max(alcohol_country$RATE_PER_CAPITA_NU))
```

Now the plot looks pretty.